### 2023-01-22 reports/report3/main.tex
Term report (report3) added

        \Large
        \textbf{FIRST INTERIM REPORT}

@@

    \textbf{Submission Date:} 05.11.2022

@@

@@

>>>>

        \Large
        \textbf{TERM REPORT}

@@

    \textbf{Submission Date:} 22.01.2023

@@

\begin{abstract}
    One of the most important aim of the modern technology is the sensing the environment. Living beings have sense organs to know the environment. They have brains to make fusion the sensor data and augment the environmental information taken from sense organs. For example, human standing in balance, we can say that they use the data coming from their eyes and ears together.

    From the point on we are studying on the sensor fusion. In this report we will briefly emphasize the legged robot usage for our final decisions. On the other hand, the concept ot the sensor fusion and the alternative methodologies to implement it correctly to serve the main purpose of sensor fusion, that is reliable odometry data extraction.
\end{abstract}

@@

\section{METHODS}

The definition of our project is explained in detail in the first interim report. In the light of the definition, we try to make the decisions and the design parameters clear. To sake of easiness we draw some paths going same destination and arguing to reach the optimal path on the alternatives. Those alternatives are listed and described below. Then we will clearly elucidate our method now.

Since the project includes plenty of areas to work those are distinct, we divide them into subtopics. Those are:

\begin{enumerate}
    \item Sensor Choosing Methods
    \item Sensor Fusion Methods
    \item Sensor Data Enhancement Methods
    \item Fusion Data Enhancement Methods
\end{enumerate}

\subsection{Sensor Choosing Methods}

Our aim in this project is to achieve high reliability by using (fusing) all the sensors we have together. However, when we want to do all these together, we foresee that the project will reach very different dimensions. Therefore, while continuing the project, we decided to select a number of sensors in order to focus on algorithm development.

We set ourselves criteria when choosing sensors. Easy to use sensors is one of our main priorities due to our job description. In addition, considering the algorithmic complexity, working with sensors with linear system response is also an important criterion. In addition, the frequency of receiving data from the sensor is of great importance. Because if the sensor data we will fuse does not come at the same frequency, we will need to add operations to our algorithm to make the sensor frequencies equal by upsampling and downsampling.

We basically divided the sensors we have into two groups as those producing RGBD data or those producing Depth Point Cloud data.

\subsubsection{Method 1: Internal Sensor Using}
The internal sensors of the legged robot include five pairs of stereo cameras that produce RGBD data, and ultrasonic sensors that produce Depth Point Cloud data when viewed cumulatively, SuperSensorySystem. In addition to these, each foot has pressure sensors and an internal IMU.

\subsubsection{Method 2: External Sensor Using}
Our external sensors are ZED Stereo camera which produces RGBD data, Velodyne 3D LiDAR which produces Depth Point Cloud data and Xbox 360 Kinect having 3D depth sensor, RGB camera, microphone array, and also tilt motor. In the future, it is also possible to add sensors such as IMU and GPS to support odometry information.

\subsubsection{Method 3: Mixed Sensor Using}
We can also use our sensors mixed. However, complexity will increase in this method and the workload to be done before the sensor fusion algorithm will become more.

\begin{itemize}
    \item \textbf{Internal Unitree Camera – Super Sensory System:} \\
          Lens Angle ~150°×170°

We have taken all these into consideration while doing the preliminary design, and together with the evaluations under the following headings, we have determined a starting point for ourselves. You can find the details under the Preliminary design heading.

% Stereo Camera Decision Method
% LiDAR Decision Method
% Internal IMU usage on legged robot
% Documentation about sensors in marketplace and we owned (engineering decisions on costs)

\subsection{Sensor Fusion Methods}

\subsubsection{Method 1: Complementary Filter}
We wanted to set out using known methods for sensor fusion. In this way, we aimed to improve the performance we want to achieve with methods that we understand fundamentally, instead of using the resources in the literature directly. In fact, we started out by examining the usages of Kalman filter and extended Kalman filter, which are quite commonly used for state estimations made from aircraft IMU sensors, in this field. We predicted that if you continue on your way by choosing the most primitive complementary filter among them, we will be able to understand more clearly that we are moving away from the target in every development we make.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.7\textwidth]{CF.png}
    \caption{Complementary Filter Example for Aircraft Attitude}
\end{figure}

\subsubsection{Method 2: Kalman Filter}
Using the Kalman filter is also a method between using the extended Kalman filter and using a complementary filter. Since we want to use it with some methods that will enhance compelemntery filtery, we have already produced our own gain function. However, we also considered the Kalman filter as a method.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.7\textwidth]{KF.png}
    \caption{Kalman Filter Example for Aircraft Attitude}
\end{figure}

\subsubsection{Method 3: Extended Kalman Filter Method}
In addition to previous methods, we also evaluated the opinion that preliminary design can speed up the process by directly using the most advanced extenden Kalman filter. However, unlike in the aircraft scenario, we could not find an example use for depth data to be suitable for our scenario~\cite{9307398}.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.7\textwidth]{EKF.png}
    \caption{Extended Kalman Filter Example for Aircraft Attitude}
\end{figure}

A comparison in which the models in which the state estimator methods are effective and the noise characteristics are given together is given in the X table, cost function comparisons for further analysis~\cite{The_cost_function_of_the_data_fusion_process_and_i}.

\begin{table}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{CostComparison.jpeg}
    \caption{Cost Function Comparison}
\end{table}

\subsection{Sensor Data Enhancement Methods}

We compare the following methods for data enhancement to following data fusion.

\subsubsection{Method 1: Only Data Rate and Range Matching}
Since the data we want to obtain is odometry, we evaluated creating parametric inputs in the methods we used while continuing the project. Because we sould think about Sensor Data Rate - State Estimation Relation. In this way, we aimed to analyze whether we could establish a relationship between the quality of the raw data and the esimated odometry data. After some tests we will briefly report our results about early sensor data enhancement for sensor fusion.

\subsubsection{Metod 2: Papoulis-Gerchberg Algorithm for Depth Point Cloud}
We are searching about Papoulis-Gerchberg algorithm effects on LiDAR and Stereo depth map results to enhance the data resolution before data fusion. An image super-resolution method, the Papoulis–Gerchberg (P–G) algorithm, on range data represented in the form of a greyscale image. However, the low convergence rate of the original P–G algorithm impedes its use for online applications~\cite{ozbay2015high, kuzucu2018enhancing}. By referencing this method we also proposed as an enhancement method.~\cite{6280128}

\section{PRELIMINARY DESIGN}

ZED stereo camera and Velodyne 3D LiDAR were chosen for the preliminary design, which are sensors that are well-documented, have a large community, and are easier to use and access resources than the internal sensors of the legged robot. This sensor will also minimize the complexity of the algorithm to be used in sensor fusion, as they offer a wide resolution and data rate band.

As the sensor fusion algorithm, it was ordered from simple to complex or from low to high performance, and it was decided to make a start with a complementary filter. Again, it was decided to use methods such as edge preservation~\cite{9307398} as an aid to the main algorithm in order to increase the sensor fusion performance.

We choose the complementary filter for fusing the depth information of two sensors. Each sensor has its own properties, advantages and disadvantages. According to these characteristics of sensors we tune our sensor fusion algorithm. But like all control systems, we first regulate our input signals to work with.

Thing we have to match in these two sensors data is viewing angle. Stereo vision camera has a viewing angle upto 110°. We simply reduced the viewing angle of the 3D lidar to match the stereo vision cameras viewing angle. But further steps we are planning to use also 360° lidar information to enhance environmental data while performing SLAM.

Raw data provided by lidar and stereo vision cameras in different formats. Stereo vision camera extracts depth frames as well as RGBD (Red Green Blue Depth) in cartesian form. 3D lidar extracts data in point cloud form which is basically in spherical form. In the other hand the resolution of stereo vision data is available in 2208x1242 at 15 fps, 1920x1080 at 30fps, 1280x720 at 60 fps, 672x376 at 100 fps. The 3D lidar works completely different and samples much less data. In order to match the sampled data from two sensors, we decided upsampling the lidar data and converting it to cartesian form. We use dynamic Papoulis–Gerchberg algorithm to augment the lidar data and then linearly interpolate this information while converting point cloud data to depth frame.

Iterative Papoulis–Gerchberg algorithm is the method used to recover the lost samples of the signal. It is mostly used for super resolution images in image processing. In a single frame manner, iterative Papoulis–Gerchberg algorithm will be enough to increase data in a single point cloud of a lidar. But our aim is continuously matching the frames, we will use the dynamic Papoulis–Gerchberg algorithm. This algorithm make use of stillness of background and continuity of movement to reduce computational effort in Papoulis–Gerchberg.

There is also difference between sample rates of two sensors. Stereo camera samples frames in range of 15 to 100 frames per second. Lidar samples point clouds by rotating its lidar continuously. Each rotation provides one whole 360° sample (30° vertical). Lidar works in range 300 rpms to 600 rpms. That provides us 5-to-10 point clouds per second. We aim to up-sample the  number of point cloud per second by again linear interpolation.

There is one more problem with the lidar. Ability of lidar is not limited by the only rotation but also sample rate of sensor. Increasing rotation speed reduces the quality of the point clouds. One part of the project is experiencing the different data rates and calculating the processing effort, transmission capabilities and response times. So we will optimize resolution and data rates in the content of this project.

\subsection*{Sensor Fusion Algorithm}

We matched the formats of the input signals to keep the sensor fusion simple. There are two depth images as input signals to get more accurate depth image.

As a preliminary design, we are planning to calculate the depth image in per pixel fashion. As mentioned before, the complementary filter is formulated as:

\begin{equation}
    \hat{x} = \hat{x}_1 \alpha + \hat{x}_2 (1 - \alpha)
\end{equation}

We will generate $\alpha$ generation function to calculate $\alpha$ for each pixel and condition. Input sources are not functioning perfect in every conditions. There are some internal and external variables that effects the accuracy of measurement. Let’s observe these variables:

\subsubsection*{Variables That Effect Lidar:}

\begin{itemize}
    \item Beam Angle:

          3D lidar scans the environment in a three-dimensional manner. Velodyne VLP-16 scans 360° horizontally and 30° vertically. In different beam angles, Lidar accuracy depends. There is an article that contains more information about beam angle effect.

          \begin{figure}[H]
              \centering
              \includegraphics[width=0.5\textwidth]{LidarAngles.jpeg}
              \caption{Velodyne LiDAR Beam Angles~\cite{VelodynePerformance}}
          \end{figure}

    \item Rotation Speed:

          Sample rate of the lidar decreases as rotation speed of the lidar increases. It effects the quality of the depth image. But increasing the rotation speed provides us approximation to real time application.

    \item Distance:

          3D lidars can work in a range that is specified in datasheet. Out of this range, lidar cannot work properly. As we think of the working principle of the lidar, the processing time is not enough to measure too close points. The Velodyne VLP-16 model has a working range between 1m to 100m.

    \item Rough and Uneven Views:

          The Lidar is using wavelengths very close to visible light to observe visible objects. In some environments like forests, rough areas, rainy and foggy weathers there are lots of indentations like branches, twigs and leaves or water drops. Each of this bits and pieces prevents the infrared laser lights coming from lidar and creates canopies. Instead of a clean point cloud, lidar gets a noisy one.

          We can measure the effect of the bits and pieces in the environment to accuracy of lidar measurement by calculating the gradient of depth frame around target pixel.

          \begin{figure}[H]
              \centering
              \includegraphics[width=0.9\textwidth]{LidarViews.png}
              \caption{Gradient of Depth Image Around Target Pixel Gives Roughness of the Environment~\cite{https://doi.org/10.48550/arxiv.2010.09076}}
          \end{figure}
\end{itemize}

\subsubsection*{Variables That Effect Stereo Vision Camera:}

\begin{itemize}
    \item Distance:

          According to viewing angles and algorithm used to extract depth information from stereo image, there is an uncertainty appears. And this uncertainty increases as distance of the pixel increases. Uncertainty reduces the confidence of the input as it increased.

          \begin{figure}[H]
              \centering
              \includegraphics[width=0.4\textwidth]{CamDistance.png}
              \caption{Stereo Camera Depth Error Calculation~\cite{DisparityCalculator, gallup2008variable}}
          \end{figure}

    \item Brightness:

          The brightness of the environment is important for stereo vision depth estimation. As considering the working principles of the stereo vision depth estimation, there are two frames taken from different angles. First step is matching each pixel of left frame with corresponding pixel in right frame. This step includes computer vision algorithms like Sum of Squared Differences (SSD) and Sum of Absolute Differences (SAD). To perform matching properly, the frames should have well distributed histograms and good contrast values. If there is the brightness of frames are too high or too low, the algorithms cannot match the corresponding pixels.

          We can calculate the effect of brightness by observing the histogram of the RGB frames.
    \item Straight and no-Contrast Views:

          Like the brightness, the content of the frame is also important. If the frame includes low contrast, straight or similar patterns, the matching algorithms won’t work properly. For example, If the frame includes a flat white wall, the algorithm can not distinguish pixels in different frames. Likewise, if the frame has a periodic pattern like damas pattern, the algorithm also not work well.

          In this case we can calculate the effect of straightness by calculating gradient around target pixel in RGB frames.
\end{itemize}

According to these parameters, we will derive confidence generator functions that will calculate a confidence value for each input. By normalizing one of the confidence values will give us $\alpha$ value.

\begin{equation}
    \alpha = \frac{C_{LIDAR}}{C_{LIDAR} + C_{STEREO}}
\end{equation}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.9\textwidth]{AlphaFunction.jpeg}
    \caption{Alpha Function for Complementary Filter Enhancement}
\end{figure}

After extracting $\alpha$ value for each pixel of depth frame, we can simply apply complementary filter.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.9\textwidth]{CompFiltBD.jpeg}
    \caption{Complementary Filter Block Diagram}
\end{figure}

\section{TEAM WORK}

Teamwork is an essential aspect of any successful project, and this project is no exception. Our team is composed of multiple members with different backgrounds, expertise, and skills, working together to achieve a common goal. To ensure effective collaboration and communication among team members, we have chosen to use Github as our primary source control and project management platform. This allows us to easily share code, track changes, and collaborate on different parts of the project.

In addition to Github, we are also using MATLAB, Python, and the Robot Operating System (ROS) environment for our project. These tools provide us with the necessary functionality and flexibility to implement and test our sensor fusion algorithm. MATLAB, in particular, is useful for prototyping and simulating our algorithm, while Python and ROS provide us with the ability to implement and test our algorithm on real-world systems.

We also keep a logbook during the project which contains our daily progress, issues encountered, and solutions implemented. This logbook acts as a reference for team members and helps keep track of the project's history and progress. Keeping a logbook is a best practice in project management, as it helps to stay organized, keep track of progress, and document any issues that may arise.

Other papers in the field also highlighted the importance of team work, communication, and collaboration in the development of sensor fusion algorithms. For example, in their paper on sensor fusion for mobile robots, authors underline that the development of a sensor fusion algorithm is a complex task that requires a multidisciplinary team with expertise in various fields such as robotics, control systems, and computer vision. They also point out that effective communication and collaboration among team members is crucial to the success of the project.

According to \cite{daniilidis2001sensor}, sensor fusion is a complex task that has been extensively studied and is widely used in various applications. In \cite{durrant2006simultaneous}, it was stated that simultaneous localization and mapping (SLAM) is an important problem in the field of robotics that requires the integration of sensor data through sensor fusion. Additionally, in \cite{silva2020edge} the authors proposed to use edge preservation methods to increase the performance of depth map filtering for stereo vision, which can be useful for our sensor fusion algorithm.

In conclusion, effective teamwork is crucial to the success of this project, and we are utilizing a variety of tools and best practices to ensure that our team can work together effectively. Our use of Github, MATLAB, Python, and the ROS environment, as well as our logbook and regular team meetings, will help us to stay organized and work effectively together to achieve our goals.

\section{COMMENTS AND CONCLUSIONS}

In conclusion, sensor fusion is an essential component of modern technology that is used to sense and augment the environment. Our proposed project aims to develop a sensor fusion algorithm for extracting depth information using data from living organisms, specifically studying the way in which zebra fish process sensor data to produce depth information. We plan to use a legged robot for the system and cloud computing for high computing power requirements. The preliminary design of the project includes using the ZED stereo camera and Velodyne 3D LiDAR as sensors and implementing a complementary filter algorithm to fuse the depth information from the two sensors, with edge preservation methods being used as an aid to the main algorithm to increase performance.

Effective teamwork and collaboration among team members are crucial to the success of this project, and we are utilizing a variety of tools and best practices such as Github, MATLAB, Python, the Robot Operating System (ROS) environment, and a logbook to ensure that our team can work together effectively. We also have been studying the sensor fusion and alternative methodologies to implement it correctly to serve the main purpose of sensor fusion, that is reliable odometry data extraction.

In future work, we plan to improve the sensor fusion algorithm by incorporating data from the zebra fish, optimize the resolution and data rates, and test the algorithm on a legged robot with cloud computing. Additionally, we plan to implement the sensor fusion algorithm on an autonomous SLAM robot to demonstrate the usefulness of the depth information in real-world applications.

The preliminary design of our sensor fusion algorithm includes the use of the ZED stereo camera and Velodyne 3D LiDAR sensors as they are well-documented, have a large community, and are easier to use and access resources than the internal sensors of the legged robot. The sensor fusion algorithm chosen for the preliminary design is a complementary filter, which is a simple and low-performance algorithm, but it can be improved by using edge preservation methods to increase the sensor fusion performance. We have also considered the different properties, advantages, and disadvantages of the sensors to optimize the sensor fusion algorithm. The stereo vision camera and 3D lidar have different viewing angles, data formats, and sample rates, thus we have to match them by reducing the viewing angle of the 3D lidar, upsampling the lidar data, converting it to cartesian form, and using dynamic Papoulis-Gerchberg algorithm to augment the lidar data. We also aim to optimize resolution and data rates in the future work to increase the quality of the point clouds and reduce the computational effort.

In addition to the previous conclusions, it is worth noting that the preliminary design of our sensor fusion algorithm was realized using the MATLAB programming environment. This choice was made because of the wide range of tools and functions available in MATLAB for image processing and sensor data manipulation, as well as its user-friendly interface and compatibility with other programming languages such as Python and C++.

By using MATLAB, we were able to effectively implement the sensor fusion algorithm, including the complementary filter, edge preservation methods, and dynamic Papoulis-Gerchberg algorithm. We also were able to test the algorithm with sample data and evaluate its performance.

It is important to note that while the preliminary design has been realized in MATLAB, we plan to implement the final version of the sensor fusion algorithm in an appropriate programming environment for the legged robot, such as ROS, to ensure the compatibility and efficiency of the algorithm in the real-world application.
